---
title: "class 11"
author: "Kalodiah Toma (A07606689)"
format: pdf
---

Here we post prcosess and inspect our modeling results from AlhpaFold2 (AF). 

My results from AF live in the folder/directory 'hivprdimer_23119'

```{r}
results_dir <- "hivprdimer_23119/"
pdb_files <-list.files(results_dir, pattern=".pdb", full.names = T)
```

We first need to align and supperpose these PDB models and we can use the 'pdbaln()' function for this:

```{r}
library(bio3d)
pdbs <- pdbaln(pdb_files, fit=TRUE, exefile="msa")
```

## The RMSD matrix

A common measure of structural dis-similarity is called RMSD (root mean square distance).

```{r}
rd <- rmsd(pdbs)
rd
```

```{r}
library(pheatmap)
rownames(rd) <- paste("m", 1:5)
colnames(rd) <- paste("m", 1:5)

pheatmap(rd)
```

Let's view these in Mol*. Here we want the fitted coords.

```{r}
xyz<-pdbfit(pdbs, outpath="fitted")
```

A full atom based fitting or supperposition did not work very well because we have multiple chains that are in different conformations. 

I want to focus my supperposition on the most invariant part (the rigid "core" if you will).

```{r}
core <- core.find(pdbs)
```

```{r}
core.inds <- core
```

```{r}
xyz<-pdbfit(pdbs, inds=core.inds, outpath="core_fitted")
```

To evaluate how good multi-chain or multi-domain models are we need to look at the PAE scores (predicted aligned error)

These are output as JSON format files. 

```{r}
pae_files <-list.files(results_dir, pattern="0.json", full.names=T)
```

```{r}
library(jsonlite)

pae1<- read_json(pae_files[1], simplifyVector = TRUE)
pae1

pae5<- read_json(pae_files[1], simplifyVector = TRUE)
pae5
```
```{r}
attributes(pae1)
pae1$max_pae
pae5$max_pae
plot.dmat(pae5$pae,
          xlab="Residue No.",
          ylab="Residue No.")
```
```{r}
plot.dmat(pae1$pae,
          xlab="Residue No.",
          ylab="Residue No.",
          zlim=c(0,30))
```
#Main points 

We can run AlphaFold on google to compute infastrcuture :)
We can read these results into R and process to help us makes sense of these models and the PAE and PDDL scores. 
