---
title: "class 10b"
author: "Kalodiah Toma (A07606689)"
format: pdf
---

Today we will finish off Lab 10 on learning about and analyzing protein sturctures.

## Comparative structure analysis of Adenylate Kinase

We will use 'bio3d' package for this analysis that starts with a single sequence. 

We will also use the 'msa' package from BioConductor. 
First we need to install the 'BiocManager' we install it the same way as other pakages with 'install.packages("")' command.  

We use 'BiocManager::install()' to install any other BioConductor package we want - like 'msa' in this case. 

```{r}
library(bio3d)
```

First we will pick a sequence of interest - I will pick "1ake-A"
```{r}
aa <- get.seq("1ake-A")
aa
```
I want to search for all related structures in the PDB database. 
```{r}
#b <-blast.pdb(aa)
```

```{r}
#hits<- plot(b)
```

```{r}
#head(b$hit.tbl)
```

```{r}
#hits$pdb.id
```
> Side-note: Lets save our results thus far so we don't have to run blast again. 

```{r}
save(hits, b, file="blast_results.Rds")
```

```{r}
load("blast_results.Rds")
```

Now we will download all these related structures from the database with 'get.pdb()'

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

##Align and superpose structures

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

```{r}
pdbs
```
## Principal Component Analysis

We could do a PCA on the coordinate data (i.e. the '$xyz' values) of all these structures.

```{r}
pc.xray <- pca(pdbs)
plot(pc.xray)
```
```{r}
dim(pdbs$xyz)
```
```{r}
plot(pc.xray, 1:2)
```

```{r}
mktrj(pc.xray, file="pca_results.pdb")
```

