---
title: "Class 5: Data Visualization with ggplot"
author: "Kalodiah Toma (PID: A07606689)"
format: pdf
---

##Using GGPLOT

To use ggplot2 we first need to install it on our computers. To do this we will use the function `install.packages()`.

Before I use any package functions I have to load them up with a `library()` call, like so:
```{r}
library(ggplot2)
ggplot(cars)
```
```{r}
head(cars)
```
There is always the "base R" graphics  system, i.e `plot()`.
To use ggplot I need to spell out at least 3 things:
-data (the stuff I want to plot as a data.frame)
-aesthetics (aes () values - how the data map to the plot, color, shape, size, etc).
-geoms (how I want things drawn, lines, heat maps, etc.)
```{r}
ggplot(cars) +
  aes(x=speed, y=dist)+
  geom_point(col="orange", size=3, alpha = 1) +
  labs(title="Speed and Stopping Distances of Cars",
       x="Speed (MPH)", 
       y="Stopping Distance (ft)",
       subtitle = "Your informative subtitle text here",
       caption="Dataset: 'cars'") +
  geom_smooth(method="lm", se=FALSE, col="black") +
  theme_bw()
```
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
nrow(genes)
colnames(genes) 
ncol(genes)
table(genes$State)
round( table(genes$State)/nrow(genes) * 100, 2 )
```
use table () table name $ column name to get column information
How many are "up"
```{r}
sum(genes$State == "up")
```
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)

table(genes$State)
```

```{r}
library(ggplot2)

p <- ggplot(genes) + 
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point() 
p
```



```{r}
#| eval: false
p + scale_color_manual(values=c("blue","gray","red")) +
    labs(title="Gene Expresion Changes Upon Drug Treatment",
         x="Control (no drug) ",
         y="Drug Treatment")
```


```{r}
# File location online
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"
gapminder <- read.delim(url)
# install.packages("dplyr")  ## un-comment to install if needed
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)

```
```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(x = gdpPercap, y = lifeExp,
                 size = pop), alpha=0.5) + 
  scale_size_area(max_size = 10)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)
ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 15)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
```


