---
title: "class 08"
author: "Kalodiah Toma (PID A07606689)"
format: pdf
---

```{r}
candy_file <- "https://github.com/fivethirtyeight/data/tree/master/candy-power-ranking"
```

```{r}
read.csv("Class 08.Rproj.csv")
candy <- read.csv("Class 08.Rproj.csv", row.names =1)
head(candy)
```
Q1. How many different candy types are in this dataset?
Q2. How many fruity candy types are in the dataset?
```{r}
nrow(candy)
```
```{r}
candy$fruity
sum(candy$fruity)
as.logical(candy$fruity)
```
Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
Q4. What is the winpercent value for “Kit Kat”?
Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
```{r}
candy["Air Heads", ]$winpercent
```
```{r}
candy["Kit Kat", ]$winpercent
candy["Tootsie Roll Snack Bars", ]$winpercent
```
```{r}
skimr::skim(candy)
```

Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
Yes, the data is reduced and dominated by the column with the most variance. 

Q7. What do you think a zero and one represent for the candy$chocolate column?
1=TRUE, there is chocolate
0=FALSE, no chocolate

Q8. Plot a histogram of winpercent values
```{r}
hist(candy$winpercent)
```


```{r}
choc.ind<-as.logical(candy$chocolate)
fruit.ind<-as.logical(candy$fruit)
choc.win<-candy[choc.ind,]$winpercent
fruit.win<-candy[fruit.ind,]$winpercent
mean(choc.win)
mean(fruit.win)
candy$winpercent[as.logical(candy$nougat)]
t.test(candy$winpercent[as.logical(candy$chocolate)], candy$winpercent[as.logical(candy$fruit)])
```

Q9. Is the distribution of winpercent values symmetrical?
No
Q10. Is the center of the distribution above or below 50%?
Below
Q11. On average is chocolate candy higher or lower ranked than fruit candy?
Yes
Q12. Is this difference statistically significant?
Yes


```{r}
x<-c(5,2,3,6)
sort(x)
sort(x, decreasing = T)
x
order(x)
x[order(x)]
```
```{r}
y<-c("D","A","E")
order(y)
y[order(y)]
```
Q13. What are the five least liked candy types in this set?
Snickers, Kit Kat, Twix, Reese's Miniatures, Reese's Peanut Butter Cup 
```{r}
head(candy[order(candy$winpercent),], n=5)
order(candy$winpercent)
tail(candy[order(candy$winpercent),], n=5)
```

Q14. What are the top 5 all time favorite candy types out of this set?
Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, Jawbusters

```{r}
library(ggplot2)
ggplot(candy)+
  aes(winpercent, rownames(candy)) +
  geom_col()
```

```{r}
ggplot(candy)+
  aes(winpercent, reorder(rownames(candy), winpercent)) +
geom_col()
```

```{r}
mycols<-rep("black", nrow(candy))
mycols[as.logical(candy$chocolate)]<-"chocolate"
mycols[as.logical(candy$bar)]<-"red"
mycols[as.logical(candy$fruit)]<-"pink"

mycols


ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=mycols) 
```

Q17. What is the worst ranked chocolate candy?
Sixlet
Q18. What is the best ranked fruity candy?
Starbust

```{r}
library(ggrepel)
```

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=mycols) + 
  geom_text_repel(col=mycols, size=3.3, max.overlaps = 5)
```
Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
Strawberry bon bons

Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
Nik L Nip

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
tail( candy[ord,c(11,12)], n=5 )
```
```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```
```{r}
library(corrplot)
```
```{r}
cij <- cor(candy)
corrplot(cij)
```
Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
chocolate and fruity
Q23. Similarly, what two variables are most positively correlated?
chocolate and bar

```{r}

pca<-prcomp(candy, scale=TRUE)
  summary(pca)
  plot(pca$x[,1:2], col=mycols, pch=16)

```
```{r}
my_data <- cbind(candy, pca$x[,1:3])
```
```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=mycols)

p
```
```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=mycols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
#library(plotly)
```
```{r}
#ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
If you are on the plus side (positive direction) of the plot you ar emore likely to be fruit, hard, and come in a pack (pluribus). Yes, this makes sense that these are on the same side of the PC1 direction. 

